### 2019-11-28 latex/sections/abstract/english.tex
latex add

this is my abstract written in the language of English I am testing thought

alas  I knew

ok so now I'm trying something new

asdf

test