### 2019-11-28 latex/sections/abstract/icelandic.tex

this is my abstract written in the language of íslensku þæööasdf