### 2019-11-28 latex/sections/discussion.tex

\chapter{Discussion}

\ifdraft{Here I will discuss the myriad of fields research like this can help
for instance for health organizations, patients, taxpayers, and pharma}
hey

\section{Summary}

\ifdraft{summarize the work}
hey

\section{Conclusion}

\ifdraft{conclude the work, discuss the significance of the work}
hey