### 2019-11-28 latex/sections/managerial.tex

%% Some elements have different names on the RU Masters rules
%% They will be annotated with RUM: "name"
\frontmatter{} % setup formatting at beginning

%\frontcover{}%%If you want to see what it looks like with the printed cover
%% TODO:  link to fill-in PDF file on RU website

\frontrequiredpages{}%% the various signaturepages and abstract
%%% WARNING:  if you get an error on the previous line, it is probably because
%%% you put a bad macro or something strange in a title, author, or abstract.

\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
  have not already done so, \LaTeX{} the \path{instructions.tex} to
  learn how to setup your document and use some of the features.  You
  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
  There is also more information on working with \LaTeX{} at
  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
  This includes common problems and fixes.

  This page will disappear in anything other than draft mode.}{}

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this to my spouse/child/pet/power animal.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
\coverchapter{Acknowledgements} 
\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}
\vspace{\baselineskip}

\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
  Note that it is important to acknowledge any funding that helped in the work}

This work was funded by \the\year~RANNIS grant ``Survey of man-eating Minke whales'' 1415550.
Additional equipment was generously donated by the Icelandic Tourism Board.

\coverchapter{Preface}

This dissertation is original work by the author, Sigurður Helgason.

\draftnote{The preface is an optional element
  explaining a little who performed what work.  See
  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
  for suggestions.

  List of publications as part of the preface is
  optional unless elements of the work have already been published.
  It should be a comprehensive list of all publications in which
  material in the thesis has appeared, preferably with references to
  sections as appropriate.  This is also a good place to state
  contribution of student and contribution of others to the work
  represented in the thesis.}

%\coverchapter{Publications}
%% RUM: Not mentioned, this was found in the CS thesis template.  
%% Maybe more applicable to PhD dissertations?
%%% Probably a duplication from before Preface became standard.

\starttables{}% setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}\clearpage%%RUM: "Table of contents"
\listoffigures{}\clearpage%%RUM: "List of figures"
\listoftables{}\clearpage%%RUM: "List of tables"

%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
DR & Diabetic Retinopathy\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
\end{tabular}

% \coverchapter{List of Symbols}%%RUM: Not mentioned
% \begin{tabular}{lll}
% Symbol &Description &Value/Units\\
% $E$ &Energy &\si{\joule}\\
% $m$ &Mass &\si{gram}\\
% $c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
% \end{tabular}