### 2019-11-28 latex/sections/results.tex

\chapter{Results}

\ifdraft{In this section you discuss any issues that came up while developing
the system.  If you found something particularly interesting,
difficult, or an important learning experience, put it here.  This is
also a good place to put additional figures and data.}
hello world