### 2020-07-11 latex/sections/background.tex
wrote to background and intro

\chapter{Background\label{cha:background}}

In this chapter we discuss the algorithms we decided to use, what some other similar 
methods exsist as well as the field in general.

\section{Environments}
\label{sec:environments}

When doing research on Artificial Intelligence it is important to select 
an environment that is a suitable abstraction for the task at hand. Environments vary 
significantly, and can be identified by their characteristic. The characteristics 
that are generally used to describe environments can be seen in Table \ref{tab:env_characteristics}. 
CITE AIMABOOK

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|p{6cm}|}
    \hline
    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
    \hline
    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
    \hline
    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
    \hline
    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
    \hline
    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
    \hline
    Static & Static, Semi-Static, Dynamic & ehh  \\
    \hline
    Discrete & Discrete, Continious & Is your environment discrete w.r.t actions (does it end) \\
    \hline
  \end{tabular}
  \caption{Charcteristics of environments}
  \label{tab:env_characteristics}
\end{table}

Categorizing environments like this gives you the power to find an environment in which 
a method works and know it can be applied to different environments with the same characteristics.
Talk about \textbf{agents} in the context of environments as the entities that act within the 
environment. A game player is an agent as well as the neural network model that exists within 
a car that is able to drive in the real world.

\section{Game Environment}

With classical Artificial Intelligence Game Environments are commonly used to validate a method, 
game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators. Game environments 
are a suitable place to apply AI as they serve an important function as an abstraction of the real 
world, for instance, a self-driving car agent who is verified to avoid driving into walls in a 
simulation is possibly safer than one who is not.

\section{Breakthrough}

The game breakthrough is a simplified version of chess, the game is set up on an $MxN$ board with 
cells like in chess, and each player starts with two rows of pawns at either end. The objective of the 
game is to move one of the pawn pieces to the opposite end of the board. A player wins if 
either they have reached the opposite end of the board, or has captured all of his opponents pawns.
The pawns differ from chess pawns in such a way that they can not move two squares on the first move and,
they can move diagonally as well as forward, this leads to the game being impossible to draw as 
pieces are always able to move. An example of an initial board in breakthrough can be seen in Figure \ref{fig:initbtboard}

\begin{figure}[]
  \centering

  \breakthrough{8/8/pppppp2/pppppp2/8/8/PPPPPP2/PPPPPP2 w - - 0 1}

  \caption{Example breakthrough board}
  \label{fig:initbtboard}
\end{figure}

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|}\hline
    \textbf{Characteristic} & \textbf{Value}\\\hline
    Observable & Fully \\
    Agents & Multi \\
    Deterministic & Deterministic\\
    Episodic & Sequential \\
    Static & Static \\
    Discrete & Discrete \\\hline
  \end{tabular}
  \caption{Categorization of Breakthrough}
  \label{tab:breakthrough_cat}
\end{table}

Categorizing Breakthrough with the characteristics described in Section \ref{sec:environments}. We end up with the description 
shown in Table \ref{tab:breakthrough_cat}. This description is identical to that of Tic-Tac-Toe, and Chess.
This description is the most common in board games where two player compete.

\subsection{Heuristics of Breakthrough}

To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance 
a very simple heuristic would be the amount of pieces each player has left. This heuristics gives us some 
insight into how well the game is progressing, but obviously there are cases where this doesn't tell us 
much, as in cases when your opponent has a single piece left that is on the row immediately before the 
row needed for him to win. No matter how many pieces you have left, this state is bad for you if you're not 
able to capture that piece. An example of such a state can be seen in Figure \ref{fig:bt_h1_bad}.

\begin{figure}[]
  \centering
  \breakthrough{8/8/pppp4/pp3P2/2p5/8/8/PPP5 w - - 0 1}
  \caption{Breakthrough board where Heuristic 1 doesn't work well}
  \label{fig:bt_h1_bad}
\end{figure}

A different heuristic would be the distance of your most advanced pawn minus your opponents most 
advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
it is favourable to move your whole team as a unit and play more defensively, so this heuristic 
is probably not one that will lead to the most wins.

\section{State-Space Search}

Traditionally methods for playing games were searching through the environment using a heuristic to
guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
as a score and give states where you win a positive score. We say that those search algorithm 
is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
This method of searching is generally extremenly inefficient as the state-spaces of game environments 
are generally extremely large, and often infinite. For instance an estimate of the state-space 
for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
$5x4$ the estimated state-space is $1310720$ many states.

This is why a good heuristic is very valuable because we can disregard all states $S'$ that 
result from doing a move $a$ in state $S$ as they will only lead to worse outcomes.

The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS),
Breadth-first Search (BFS), Alpha* Search (A-Star), and Monte-Carlo Tree Search (MCTS).

More modernly these search methods have been amplified by Machine Learning, in such a way that 
we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
learning model to learn a function that takes in a state and returns an evaluation of that state.
This can lead to a significant time reduction as we do not need to simulate a whole game from a 
state to recieve its evaluation we rather receive the evaluation from the model.

\subsection{Monte-Carlo Tree Search}

\label{sec:mcts}

In the algorithm Monte-Carlo Tree Search there is an agent exists
within some environment. Where each node in the environment represents a state-action pair of the environment, by
state-action pair what is meant it is some state and the action that brought the agent to
that state. This pair should be unique within the environment. We will discuss these concepts within game-environments
and therefore we might say player instead of agent and game instead of environment.

MCTS is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Backpropagation. They happen sequentially
and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.

In MCTS there is a tree representing the game-environment. This tree consists of nodes $n_i$ where $i$ represents the point in time
of the node, for example $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of
the states representing a position where there is either a draw or one player has won the match. Each of the nodes has $4$ values,
$s$, $a$, $Q$, and $N$. These values represent these items, $s$ is state of the environment, $a$ is the action that brought
the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game),
and $N$ the amount of times the MCTS algorithm has visited this node. The values $s$ and $a$ uniquely identify a position
in the environment and are often called state-action pairs.

The MCTS algorithms four phases
\begin{enumerate}
  \item Selection
  \item Expansion
  \item Simulation/Rollout
  \item Backpropagation
\end{enumerate}

\begin{equation} \label{UCT_formula}
  \text{Child UCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
\end{equation}

\subsection*{Selection}
During the selection phase a node $(s,a)$ within the tree which has not yet been expanded is found.
This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
(initially the root of the tree) we select the child with the highest UCT value. Repeatedly until an unexpanded
node is found. This process is done to balance the amount of exploration vs exploitation of nodes in the
tree.

\subsection*{Expansion}
Then the expansion phase expands the node generating all of $(s,a)$'s children $(s',a')$ are generated
and connected to $(s,a)$. This is done by applying all actions $a'$ to $(s,a)$. These children are the
product of applying each action $a'$ to $s$ in the parent node.

\subsection*{Rollout}
Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal we mean a
state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
context of this paper we only return (+1 white wins, -1 black wins, 0 draw).

\subsection*{Backpropagation}
The result from the terminal node is then propagated up through the path taken by selection $(s,a)$ up
to the root of the tree, updating the $Q_{(s,a)}$ values of each node $(s,a)$.

When training a neural network the UCT formula is modified slightly to prefer selecting nodes
that the neural network values highly by introducing a second scalar to the formula $f(s,a) = (p,v)$,
the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
is modified to instead of doing rollout/simulation to recieve a reward the predicted value $v$ from
the neural network is used instead.

\begin{equation} \label{PUCT_formula}
  \text{Child PUCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * P((s,a)) * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
\end{equation}

\section{Machine Learning}

Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
extrapolate some underlying function of that data. This some task like predicting the 
rise and fall of some stock, what the weather will be in a week, and whether an image 
is an image of a dog or a cat. Many datastructures and algorithms are used to achieve this 
goal, but recently the field of ANN/DNN's has been the standard for achieveing the best
results.

\subsection{Reinforcement Learning}

Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
That is, the notion of input/output values changes, the model uses itself to generate input
values by acting in an environment. The environment then returns some result, that could be
losing in a game, making a correct prediction, or any number of outcomes. The result is then
propagated through the model allowing it to improve with this new information.
% When doing moving it's parameters in the direction of the recieved output
% from the environment.

Many algorithms are popular in reinforcement learning, for instance Q-learning(CITE ME), CARLA (CITE ME), 
and many others.

\subsection{Neural Networks}

Neural networks (NN) are popular methods within a subfield of ML which is called Deep Learning (DL). 
NN's are created to resemble how the human brain functions. In the brain we have neurons which 
when they get a signal they apply some function to them and if the resulting signal is high 
enough, they fire to the next neuron. This is how it is done in the neural network model as well.
There we have neurons that when they get some input, generally a vector of numbers. The neuron 
takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
of neurons is reached. At that point we have a value that the neural network corresponds to the 
input value. This value can be a binary classification (cat or dog image), a regression value (the 
value of a property), or any number of outputs. It can then be said that a neural network is doing 
a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.

Neural networks are machine learning models that take in as input anything that is numerical 
and they apply differentiatable function to that input s.t. they end with some output. This 
output is then compared to the expected output the difference between these outputs is called 
a loss. A loss is backpropagated through the neural network, and each function is derived in 
order to find its slope. We can then modify the weights of the neural network in the direction 
of the correct output. Leading to a function approximation of this function
$f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$

\section{Explainable Artificial Intelligence (XAI)}

The field of XAI research is still very far behind it's counterpart AI research,
within XAI two fields are concidered the largest that is Model Interpretability
and Model Explainability.

\subsection{Model Interpretability}

\subsubsection{Saliency Maps}
Within XAI many methods have been developed to try to evaluate
ANN's, these methods are often referred as model interpretability methods.
In the field image recognition there has been a lot of work examining which
pixels of an image the model deems important. Arguably the most popular
method for this is Saliency Maps (CITE ME). There the pixel values the model
deems important are colored in s.t. a human can examine the image and get a
sense of what portions of the image are important to the model, an example of
a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.

\begin{figure}[]
  \centering
  \includegraphics[width=.5\textwidth]{graphics/dog_saliency}
  \caption{Example of a models saliency map for an image of a dog}
  \label{fig:dog_saliency}
\end{figure}

While this method is understandable in the context of image recognition it
lacks severely when your input is not an image.

\subsubsection{Shapley Values}

Methods for explaining models that aren't image recognition models include
shapley values. There the input is examined against it's output, then iteratively
input values are selected to be fixed. Then the other input values are varied and
an average change in prediction is calculated. With this the shapley value can be
estimated for the fixed input value. This is done to examine which input values have
the strongest link to the output value. Shapley values on a dataset can give insights
on which input values the model deems important.

\subsubsection{Concept Activation Vectors}

A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
a neural network giving a much more human insight into a prediction. Using Concept
Activation Vectors (CAV) a directional derivative for a given input can be examined
with respect to some HLC's. For example, when a human looks at an image of an animal
and is supposed to decide whether the image is of a horse or a zebra, an intuitive
approach would be to check whether the animal has stipes, or is both white and black.
That method of determining if a horse is a zebra could then be called a higher-level
concept, and if we're able to gather if a nerual network uses this strategy for prediction 
we have a deeper understanding of its underlying structure. Leading to an explanation of 
the result.

CAV's in neural networks are just binary classifiers of the same dataset used to train
a neural network. The

\subsection{Model Explainability}

Model explainability within the context of neural networks isn't possible today. Model
explainability referrs to firstly considering some input and output from a model. Then
afterwards the model is examined to determine exactly what led to the predicted output.
This concept is simple when we're working with Decision Trees. A decision tree is a tree
whose nodes are representative of an input value and at every node a branch is selected
based on the value of the input value. It is therefore easy to see how to examine the tree
to explain the output. We just follow the branches in the tree. That being said the branches
are created by algorithms like ID3 which construct branches based on the initial dataset used
to construct the tree, but again ID3 follows simple statistics and can be explained properly.

When we talk about neural networks this process is much more difficult, the underlying nodes
are generally in the thousands, the different layers of the neural network varies in the operations
it applies to the input value and such while travelling through the neural network the modified
value becomes far removed from the initial input value to the eyes of the reader. That being said,
while the possibility of completely monitoring the training process and completely monitoring the
evaluation process is truly possible it is not feasible. And secondly the process of seeing an
input and it's corresponding output will not be of any value if one were to consider the process
of prediction.