### 2020-07-17 latex/sections/methods.tex

\chapter{Methods\label{cha:methods}}

In this chapter we train a neural network to play the game breakthrough described in 
Chapter \ref{cha:background}. The training process uses MCTS described in Section \ref{sec:mcts} 
to guide it's training.

The training algorithm is a reinforcement learning based self-play algorithm based on the work 
done in AlphaZero (CITE ME).

\subsection{Neural network Architecture}

The neural network architecture we opted to use was a single convolutional layer with a ReLU 
activation function, followed by 5 residual layers, and lastly a split policy/value head.
A figure describing the architecture can be seen in Figure \ref{fig:nnarch}.

\begin{figure}[]
    \centering

    \includegraphics[width=0.65\textwidth]{graphics/test}

    \caption{Neural Network Architecture}
    \label{fig:nnarch}
\end{figure}

\subsection{Policy head}

The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move in the space (diagonally both directions, forward and backwards). The vector is then masked s.t. the values representing moves that can not be played on the board are given the value of $0$.

\subsection{Self-play}

To train the neural network to play Breakthrough we initialize two neural networks $N_1$ and $N_2$ 
with random weights. Then we let $N_1$ play against itself using MCTS to direct its training. 
While playing against itself the neural network gathers data for it to train with.

The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
MCTS for $N$ iterations, $\tau$ is the reward for the whole episode, $1$ if white wins,
$-1$ if black wins, and $0$ on a tied game, $p$ is the policy vector
from the neural network, and $v$ is the predicted reward of the game by the neural network.

\subsection{Loss Function}

The data collected is backpropagated through the NN moving the weights to the direction of this 
loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.