### 2020-12-16 latex/sections/abstract/english.tex
yngvi+james comments fixed

Deep neural networks have exceeded expectations in many areas, they can generate powerful and accurate predictions of the real world. This is going to be a powerful tool for the future, and we require a method of examining them to understand why and how they come to a specific prediction to alleviate trust issues. This paper examines the strategies of a deep neural network that plays the game Breakthrough uses. We explain the process of creating a neural network that can achieve success in playing board games only from playing against itself, without any human input. We outline the process of creating a model to examine the internal representation of the neural network of the game state to evaluate whether it both recognizes and uses strategies applied by humans that play the game. These strategies are compared against each other to infer which is best, and which the neural network uses the most. We show that the neural network emphasizes strategies that are considered the most successful handcrafted strategies for the game Breakthrough. Lastly, we examine a trained neural network with a suite of simple to complex strategies to understand which is used most.