### 2020-12-16 latex/sections/managerial.tex

%% Some elements have different names on the RU Masters rules
%% They will be annotated with RUM: "name"
\frontmatter{} % setup formatting at beginning

%\frontcover{}%%If you want to see what it looks like with the printed cover
%% TODO:  link to fill-in PDF file on RU website

\frontrequiredpages{}%% the various signaturepages and abstract
%%% WARNING:  if you get an error on the previous line, it is probably because
%%% you put a bad macro or something strange in a title, author, or abstract.

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family, who while not understanding most of what I do,
  always support me and the friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly, my girlfriend Hulda, who always helps me get 
  through the tough times, and makes the good times even better.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
\coverchapter{Acknowledgements} 
\begin{quotation}
Throughout the writing of my thesis I always had support and assistance from the people around me.

I would first like to thank my supervisor, Yngvi Björnsson, his knowledge in this field is vast and he is always ready to share invaluable wisdom when it's needed. Your critical feedback and consistent positivity has pushed me and my work to a higher level.

Secondly, I would like to thank my colleagues, who provided stimulating conversations on the topic and a helpful eye from a different perspective every time I needed it.

\end{quotation}
\vspace{\baselineskip}

%\coverchapter{Publications}
%% RUM: Not mentioned, this was found in the CS thesis template.  
%% Maybe more applicable to PhD dissertations?
%%% Probably a duplication from before Preface became standard.

\starttables{}% setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}\clearpage%%RUM: "Table of contents"
\listoffigures{}\clearpage%%RUM: "List of figures"
\listoftables{}\clearpage%%RUM: "List of tables"

%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
ANN & Artificial Neural Network\\
DNN & Deep Neural Network\\
MCTS & Monte-Carlo Tree Search\\
CAV & Concept Activation Vector\\
DL & Deep Learning\\
MI & Model Interpretability\\
ME & Model Explainability\\

\end{tabular}
\overfullrule=0pt